---
title: "Presenting Raster Bricks"
author: "Kassel Hingee"
date: "12/12/2019"
output: 
  html_document: 
    keep_md: yes
    toc: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, root.dir = "..")
```

A document for storing / demonstrating methods for plotting RasterBricks with many layers.

```{r dataimport}
invisible(lapply(c("raster", "maptools", "rgdal", "ncdf4", "lubridate", "sf"),
                 library, character.only = TRUE))
invisible(lapply(paste0("./functions/", list.files("./functions")), source))
sws_sites <- readRDS("./private/data/clean/sws_sites.rds")
points <- sws_sites_2_spdf(sws_sites)
b <- brick_fmc(points, 2001)
```

## Using ggplot2
```{r libraryprep}
library(ggplot2)
library(rasterVis)
library(viridis)
library(ggrepel)
```

Facet Grids using two different methods:
```{r facetgrids}
gplot(subset(b, 1:2)) +
  geom_tile(aes(fill = value)) +
  facet_grid(~ variable) + 
  scale_fill_viridis() +
  coord_fixed()
out <- levelplot(subset(b, 1:2))
```


Animate
```{r animate}
library(gganimate)
anim <- gplot(subset(b, 1:10)) +
  geom_tile(aes(fill = value)) +
  scale_fill_viridis() +
  coord_fixed() +
  transition_manual(variable) +
  ggtitle("Fuel Moisture Content at {current_frame}") +
  xlab("Longitude") + ylab("Latitude")
animate(anim, duration = 5)
```

Export an animation
```{r exportanim}
exportdir <- tempdir()
anim_save(filename = "testanimsave.gif", path = exportdir)
```

## Spatial Context Information
I think Open Street Maps can not be downloaded direclty using an R package (see https://github.com/dkahle/ggmap/issues/117).
Google maps requires an API key to a google account and possibly for said google account to have a credit card (see help for `register_google` in package ggmap.
Stamen Maps are a possibiity to download using `ggmap`, however it appears `ggmap` creates raster objects for these.

### Manual Download of OSM Data
Another option is to download OSM data manually and then import into R.
Downloaded a map by going to https://www.openstreetmap.org/export#map=8/-35.326/148.189 and selecting download via 'Overpass API'.
```{r OSM}
comphome <- "C://UserData/hingeek"
#st_layers(paste0(comphome, "/", "map.osm"))
lns <- st_read(paste0(comphome, "/", "map.osm"), layer = "lines")
lns <- st_crop(lns, b)
unique(lns$highway)
ggplot() +
  geom_sf(aes(col = highway), data = lns[!is.na(lns$highway), ]) +
  ggtitle("Roads")
ggplot() +
  geom_sf(aes(col = waterway), data = lns[!is.na(lns$waterway), ]) +
  ggtitle("Waterways")

lns %>%
  dplyr::mutate(
    type = dplyr::case_when (
      highway %in% c("secondary", "secondary_link") ~ "largeroad",
      waterway == "river" ~ "river"
    )
  ) %>%
  dplyr::filter(!is.na(type)) %>%
  ggplot() + 
  geom_sf(aes(col = type)) 
```

```{r combinewraster}
bigftrs <- lns %>%
  dplyr::filter(waterway == "river" | highway == "secondary") %>%
  dplyr::mutate(type = paste(highway, waterway))

gplot(subset(b, 1)) +
  geom_tile(aes(fill = value)) +
  scale_fill_viridis() +
  #do not inherit aes in following because don't want to a fill colour given by value
  geom_sf(data = bigftrs, inherit.aes = FALSE, stat = "sf", aes(col = "type")) + 
  coord_sf() 
```

### Data Supplied by Geoscience Australia
A list of the relevant data is [here](http://www.ga.gov.au/scientific-topics/national-location-information/topographic-maps-data/digital-topographic-data).
Selecting GEODATA TOPO 250K Series 3 opens a new page, defaulting to esri Personal database. One of the 'related' links on the side is for the same data in shape file format.
That webpage currently has address [https://ecat.ga.gov.au/geonetwork/srv/eng/catalog.search#/metadata/64058].

According to metadata documents with this data, the projection is latitude and longitude, and the datum is GDA94.
The datum GDA94 uses GRS80 as the ellipsoid ([Geocentric Datum of Australia 2020 Technical Manual](https://www.icsm.gov.au/sites/default/files/GDA2020TechnicalManualV1.1.1.pdf) available from [here](https://www.icsm.gov.au/datum/geocentric-datum-australia-1994-gda94). I'm not sure what other things the datum specifies beyond the ellipsoid.
The CRS information extracted from the shape files agrees - they give a proj4string of "+proj=longlat +ellps=GRS80 +no_defs".


```{r GAdata}
railways <- st_read("C:/UserData/hingeek/GA_TOPO_250K_Series_3_shp/Vector_data/Transport/railways.shp")
railways <- railways %>%
  st_transform(st_crs(b)) %>%
  st_crop(b) 

#loading roads below takes a minute or two
roads <- st_read("C:/UserData/hingeek/GA_TOPO_250K_Series_3_shp/Vector_data/Transport/roads.shp")
roads <- roads %>%
  st_transform(st_crs(b)) %>%
  st_crop(b) 
roadsplt <- roads %>%
  ggplot() + 
  geom_sf(aes(col = CLASS))

waterareas <- st_read("C:/UserData/hingeek/GA_TOPO_250K_Series_3_shp/Vector_data/Hydrography/watercourseareas.shp")
waterareas <- waterareas %>%
  st_transform(st_crs(b)) %>%
  st_crop(b) 
waplt <- waterareas %>%
  ggplot() + 
  geom_sf(aes(col = HIERARCHY))

waterlines <- st_read("C:/UserData/hingeek/GA_TOPO_250K_Series_3_shp/Vector_data/Hydrography/watercourselines.shp")
waterlines <- waterlines %>%
  st_transform(st_crs(b)) %>%
  st_crop(b) 
wlplt <- waterlines %>%
  dplyr::filter(HIERARCHY == "Major") %>%
  ggplot() + 
  geom_sf()

builtupareas <- st_read("C:/UserData/hingeek/GA_TOPO_250K_Series_3_shp/Vector_data/Habitation/builtupareas.shp")
builtupareas <- cbind(builtupareas, cen = builtupareas %>% st_centroid() %>% st_coordinates)
builtupareas <- builtupareas %>%
  st_transform(st_crs(b)) %>%
  st_crop(b) 
```

```{r plainbackground}
ggplot() +
  geom_sf(data = waterlines %>% dplyr::filter(HIERARCHY == "Major"), stat = "sf", col = "green") + 
  geom_sf(data = roads %>% dplyr::filter(CLASS %in% c("Principal Road", "Dual Carriageway")), inherit.aes = FALSE, stat = "sf", col = "red", lwd = 1) + 
  geom_sf(data = builtupareas %>% dplyr::filter(SHAPE_Area > quantile(builtupareas$SHAPE_Area, 0.8)),
          inherit.aes = FALSE, stat = "sf", fill = "black", col = "black", lwd = 1) +
  coord_sf() +
  geom_text_repel(aes(x = cen.X, y = cen.Y, label = NAME),
        data = builtupareas %>% dplyr::filter(SHAPE_Area > quantile(builtupareas$SHAPE_Area, 0.8)),
        inherit.aes = FALSE, 
        nudge_y = 0.1,
        col = "black",
        size = 3) +
  labs(x = "", y ="")
```


```{r GAontopofraster}
gplot(subset(b, 1)) +
  geom_tile(aes(fill = value)) +
  scale_fill_viridis() +
  #do not inherit aes in following because don't want to a fill colour given by value
  geom_sf(data = waterareas, inherit.aes = FALSE, stat = "sf", col = "green") + 
  geom_sf(data = waterlines %>% dplyr::filter(HIERARCHY == "Major"), inherit.aes = FALSE, stat = "sf", col = "green") + 
  geom_sf(data = roads %>% dplyr::filter(CLASS == "Principal Road"), inherit.aes = FALSE, stat = "sf", col = "black", lwd = 1) + 
  coord_sf()
```

Combine into an animation
```{r GA_raster_animate}
gplot(subset(b, 1:5)) +
  geom_tile(aes(fill = value)) +
  scale_fill_viridis() +
  #do not inherit aes in following because don't want to a fill colour given by value
  geom_sf(data = waterareas, inherit.aes = FALSE, stat = "sf", col = "green") + 
  geom_sf(data = waterlines %>% dplyr::filter(HIERARCHY == "Major"), inherit.aes = FALSE, stat = "sf", col = "green") + 
  geom_sf(data = roads %>% dplyr::filter(CLASS == "Principal Road"), inherit.aes = FALSE, stat = "sf", col = "black", lwd = 1) + 
  coord_sf() +
  transition_manual(variable) +
  ggtitle("Fuel Moisture Content at {current_frame}") +
  xlab("Longitude") + ylab("Latitude")
```


## SWS Points and GA Context
```{r swspoints}
swspoints <- sws_sites_2_sf(readRDS("./private/data/clean/sws_sites.rds"))
```

```{r animatewithpoints_GA}
gplot(subset(b, 1:5)) +
  geom_tile(aes(fill = value)) +
  scale_fill_viridis() +
  #do not inherit aes in following because don't want to a fill colour given by value
  geom_sf(data = waterareas, inherit.aes = FALSE, stat = "sf", col = "green") + 
  geom_sf(data = waterlines %>% dplyr::filter(HIERARCHY == "Major"), inherit.aes = FALSE, stat = "sf", col = "green") + 
  geom_sf(data = roads %>% dplyr::filter(CLASS == "Principal Road"), inherit.aes = FALSE, stat = "sf", col = "black", lwd = 1) + 
  geom_sf(data = swspoints, inherit.aes = FALSE, stat = "sf", col = "red") + 
  coord_sf() +
  transition_manual(variable) +
  ggtitle("Fuel Moisture Content at {current_frame}") +
  xlab("Longitude") + ylab("Latitude")
```


## Use Pre-Subsetted GA Data
```{r presubsetted_GA}
majorfeatures <- readRDS("./private/data/GA_principalroads_majorrivers_railsways.rds")  %>% 
  st_transform(st_crs(b)) %>%
  st_crop(b)
builtupareas <- readRDS("./private/data/GA_builtupareas.rds")  %>% 
  st_transform(st_crs(b)) %>%
  st_crop(b)

gplot(subset(b, 1:5)) +
  geom_tile(aes(fill = value)) +
  scale_fill_viridis() +
  #do not inherit aes in following because don't want to a fill colour given by value
  geom_sf(aes(col = FEATTYPE), data = majorfeatures, inherit.aes = FALSE, stat = "sf") +
  geom_sf(data = swspoints, inherit.aes = FALSE, stat = "sf", col = "red") + 
  geom_sf(data = builtupareas %>% dplyr::filter(SHAPE_Area > quantile(builtupareas$SHAPE_Area, 0.8)),
          inherit.aes = FALSE, stat = "sf", fill = "black", col = "black", lwd = 1) +
  coord_sf() +
  geom_text_repel(aes(x = cen.X, y = cen.Y, label = NAME),
        data = builtupareas %>% dplyr::filter(SHAPE_Area > quantile(builtupareas$SHAPE_Area, 0.8)),
        inherit.aes = FALSE, 
        nudge_y = 0.1,
        col = "black",
        size = 3) +
  transition_manual(variable) +
  ggtitle("Fuel Moisture Content at {current_frame}") +
  xlab("Longitude") + ylab("Latitude")
```

And to avoid use of aes for the context (if aes needed for something else):
```{r GAcontext_aesavoidance}
contextlyrs <- list(
  geom_sf(data = majorfeatures %>% dplyr::filter(FEATTYPE %in% c("Major Road")),
          inherit.aes = FALSE, col = "grey"), 
  # geom_sf(data = majorfeatures %>% dplyr::filter(FEATTYPE %in% c("Major Watercourse")),
  #         inherit.aes = FALSE, lty = "dotted", col = "grey"),
  geom_sf(data = builtupareas %>% dplyr::filter(SHAPE_Area > quantile(builtupareas$SHAPE_Area, 0.8)),
          inherit.aes = FALSE, fill = "grey", col = "grey", lwd = 1),
  geom_text_repel(aes(x = cen.X, y = cen.Y,
                      label = NAME), #to title case: tools::toTitleCase(tolower(as.character(NAME)))),
                  data = builtupareas %>% dplyr::filter(SHAPE_Area > quantile(builtupareas$SHAPE_Area, 0.8)),
                  inherit.aes = FALSE,
                  nudge_y = 0.1,
                  col = "grey",
                  size = 3)
)

gplot(subset(b, 1)) +
  geom_tile(aes(fill = value)) +
  scale_fill_viridis() +
  contextlyrs +
  geom_sf(data = swspoints, inherit.aes = FALSE, stat = "sf", col = "red") + 
  coord_sf() +
  ggtitle("Fuel Moisture Content") +
  xlab("Longitude") + ylab("Latitude")
```


## With Leaflet and NSW Aerial Imagery
```{r leafleteg}
library(leaflet)
your.map <- leaflet(swspoints) %>% 
  addTiles() %>%
  ## NSW aerial photography maps aren't useful for this region
  addProviderTiles("Esri.WorldImagery") %>%
  addWMSTiles("https://maps.six.nsw.gov.au/arcgis/services/public/NSW_Imagery/MapServer/WMSServer",
              options = WMSTileOptions(format = "image/png", transparent = F),
              layers = "0") %>%
  addCircleMarkers()
```