---
title: "Visualising SWS Bird Abundance Data"
author: "Kassel Hingee"
date: "14/02/2020"
output: 
  html_document: 
    fig_caption: yes
    fig_width: 10
    toc: yes
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%")
```

```{r importlibs, echo = FALSE}
out <- lapply(c("sf", "tsibble", 'lubridate', "viridis",
                'ggplot2', 'tidyr', 'grid', 'gridExtra', 
                'feasts', 'dplyr', 'gtable', 'fable',
                'mgcv'),
       library, character.only = TRUE)
out <- lapply(paste0("./functions/", list.files("./functions/")), source)
```

## Import Data
```{r importdata}
bird_richness <- readRDS("./private/data/clean/sws_bird_richness.rds") # contains all bird spp
birds <- readRDS("./private/data/clean/sws_birds.rds") # contains only common bird spp
stopifnot(all.equal(bird_richness[, names(birds)], birds))

sites_rs <- readRDS("./private/data/clean/sws_sites_rs.rds") %>%
    mutate(farm = substr(SiteCode, 1, 4),
         sitenum = factor(as.integer(substr(SiteCode, 5, 5))))
sites_rs$log_plus_one_woody_cover <- log(sites_rs$woody_cover + 1)
traits <- readRDS("./private/data/clean/sws_traits.rds")

obsa <- cbind(sites_rs, bird_richness) 
```

```{r makepresenceabsence}
# make binary
for(i in 1:ncol(birds)){
  birds[which(birds[, i] > 0), i] <- 1
}
for(i in 1:ncol(bird_richness)){
  bird_richness[which(bird_richness[, i] > 0), i] <- 1
}
obsb <- cbind(sites_rs, bird_richness)
```

## Non-Graphical Summary of Traits
```{r traits}
traits <- traits %>%
  mutate_at(c("family", "habitat", "diet", "substrate", "movements", "nest"),
            as.factor)
traits %>%  summary()
```


## Visualise Abundance
```{r vis_obsa_magpies}
obsa %>%
  #filter(farm == "BELL") %>%
  ggplot() +
  facet_wrap(vars(farm)) +
  geom_point(position = "jitter", #position_jitter(width = 1E6, height = 0),
    aes(x = SurveyDate, y = `Australian Magpie`)
             ) +
  scale_y_log10() +
  ggtitle("Australian Magpie Numbers Observed")
```

Numbers of detected Austaralian Magpie appears to have roughly constant density.

```{r vis_obsa_apostlebird}
obsa %>%
  #filter(farm == "BELL") %>%
  ggplot() +
  facet_wrap(vars(farm)) +
  geom_point(position = "jitter", #position_jitter(width = 1E6, height = 0),
    aes(x = SurveyDate, y = `Apostlebird`)
             ) +
  scale_y_log10() +
  ggtitle("Apostlebird Numbers Observed")
```


```{r vis_obsa_apostlebird_b}
obsa %>%
  #filter(farm == "BELL") %>%
  ggplot() +
  # facet_wrap(vars(farm)) +
  geom_point(aes(x = farm, y = `Apostlebird`, col = year(SurveyDate))
             ) +
  scale_color_viridis(name = "Survey Year") +
  scale_y_log10() +
  ggtitle("Apostlebird Numbers Observed")
```

That the same farms have a variety of colours suggests that the things that are changing in time are not an enormous factor on Apostlebirds.


```{r vis_obsa_all}
p <- obsa %>%
  pivot_longer(cols = names(bird_richness), names_to = "species", values_to = "numbers") %>%
  #filter(farm == "BELL") %>%
  ggplot() +
  # facet_wrap(vars(species)) +
  # ggforce::facet_wrap_paginate(vars(species), nrow = 4, ncol = 13) +
  geom_point(position = position_jitter(height = 0.2),
    aes(x = farm, y = numbers, col = year(SurveyDate))
             ) +
  scale_color_viridis(name = "Survey Year") +
  scale_y_log10() +
  theme(strip.text.x = element_text(size = 8)) +
  ggtitle("Observed Bird Numbers")
p + ggforce::facet_wrap_paginate(vars(species), nrow = 4, ncol = 13, page = 1)
p + ggforce::facet_wrap_paginate(vars(species), nrow = 4, ncol = 13, page = 2)
p + ggforce::facet_wrap_paginate(vars(species), nrow = 4, ncol = 13, page = 3)
```


Looks like the Noisy Friarbird was widely around in early years and later years, but not years around 2010.
There are a number of species that were only observed much less in later years (some of this could be confounded with any farms/sites that dropped out of the study).

## Visualise Presence/Absence Binary Simplification
```{r vis_obsb_magpies}
obsb %>%
  ggplot() +
  facet_wrap(vars(farm)) +
  geom_point(position = position_jitter(width = 0.5, height = 0.05),
    aes(x = SurveyDate, y = `Australian Magpie`)
             ) +
  # theme(axis.text.x = element_text(angle = 90)) +
  theme(strip.text.x = element_text(size = 4)) +
  ggtitle("Australian Magpie Presence Absence")
```

```{r vis_obsb_magpies_spatial}
obsb %>%
  ggplot() +
  geom_point(position = position_jitter(width = 0.5, height = 0.05),
    aes(x = SurveyDate, y = `Australian Magpie`)
             ) +
  # theme(axis.text.x = element_text(angle = 90)) +
  theme(strip.text.x = element_text(size = 4)) +
  ggtitle("Australian Magpie Presence Absence")
obsb %>%
  # filter(!(is.na(longitude))) %>%
  # filter(!(is.na(latitude))) %>%
  # sf::st_as_sf(coords = c("longitude", "latitude"), crs = "+proj=longlat +datum=WGS84") %>%
  mutate(`Australian Magpie` = as.logical(`Australian Magpie`)) %>%
  ggplot() +
  geom_point(position = position_jitter(width = 0.01, height = 0.01),
    aes(x = longitude, y = latitude,
                 col = year(SurveyDate),
        alpha = `Australian Magpie`,
        shape = `Australian Magpie`)) +
  coord_fixed() +
  scale_shape_manual(values = c("TRUE" = 17, "FALSE" = 3)) +
  scale_color_viridis() +
  ggtitle("Australian Magpie Presence Absence",
          subtitle = "Locations jittered")
```

```{r vis_obsb_apostlebird}
obsb %>%
  mutate(`Apostlebird` = as.logical(`Apostlebird`)) %>%
  ggplot() +
  geom_point(position = position_jitter(width = 0.01, height = 0.01),
    aes(x = longitude, y = latitude,
                 col = year(SurveyDate),
        alpha = `Apostlebird`,
        shape = `Apostlebird`)) +
  coord_fixed() +
  scale_shape_manual(values = c("TRUE" = 17, "FALSE" = 3)) +
  scale_color_viridis() +
  ggtitle("Apostlebird Presence Absence",
          subtitle = "Locations jittered")
```


```{r vis_obsb_all}
p <- obsb %>%
  pivot_longer(cols = names(bird_richness), names_to = "species", values_to = "detected") %>%
  #filter(farm == "BELL") %>%
  ggplot() +
  # facet_wrap(vars(species)) +
  # ggforce::facet_wrap_paginate(vars(species), nrow = 4, ncol = 13) +
  geom_point(position = position_jitter(height = 0.05),
    aes(x = farm, y = detected, col = year(SurveyDate))
             ) +
  scale_color_viridis(name = "Survey Year") +
  theme(strip.text.x = element_text(size = 8)) +
  ggtitle("Detected Birds")
p + ggforce::facet_wrap_paginate(vars(species), nrow = 4, ncol = 13, page = 1)
p + ggforce::facet_wrap_paginate(vars(species), nrow = 4, ncol = 13, page = 2)
p + ggforce::facet_wrap_paginate(vars(species), nrow = 4, ncol = 13, page = 3)


p <- obsb %>%
  pivot_longer(cols = names(bird_richness), names_to = "species", values_to = "detected") %>%
  filter(!is.na(detected)) %>%
  mutate(detected = as.logical(detected)) %>%
  ggplot() +
  geom_point(position = position_jitter(width = 0.01, height = 0.01),
    aes(x = longitude, y = latitude,
                 col = year(SurveyDate)),
        shape = 3,
    size = 2) +
  coord_fixed() +
  scale_color_viridis() +
  theme(strip.text.x = element_text(size = 6)) +
  ggtitle("Bird Presence",
          subtitle = "Locations jittered")
p + ggforce::facet_wrap_paginate(vars(species, detected), nrow = 4, ncol = 12, page = 1) 
p + ggforce::facet_wrap_paginate(vars(species, detected), nrow = 4, ncol = 12, page = 2) 
p + ggforce::facet_wrap_paginate(vars(species, detected), nrow = 4, ncol = 12, page = 3) 
p + ggforce::facet_wrap_paginate(vars(species, detected), nrow = 4, ncol = 12, page = 4) 
p + ggforce::facet_wrap_paginate(vars(species, detected), nrow = 4, ncol = 12, page = 5) 
p + ggforce::facet_wrap_paginate(vars(species, detected), nrow = 4, ncol = 12, page = 6) 
p + ggforce::facet_wrap_paginate(vars(species, detected), nrow = 4, ncol = 12, page = 7) 
```

The above plots are not overly helpful. Some species are clustered into a subregion. For example the Little Lorikeet, the White-Faced Honey Eater, the Dollarbird and Horsfield’s Bushlark.
*are these clusters environmental (correlation purely due to predictors)? or due to colonisation costs (conditional spatial correlation)?*

The species with many many detections do not have much observable spatial patttern.
There are also many species with very few detections, which means no spatial patterns could be observed.

## Covariates in Data
```{r covariatenames}
covarnames <- names(sites_rs)
summary(obsb[, covarnames])
```

### Growth Type
```{r vis_obsa_GrowthType}
p <- obsa %>%
  pivot_longer(cols = names(bird_richness), names_to = "species", values_to = "numbers") %>%
  #filter(farm == "BELL") %>%
  ggplot() +
  # facet_wrap(vars(species)) +
  # ggforce::facet_wrap_paginate(vars(species), nrow = 4, ncol = 13) +
  geom_point(position = position_jitter(height = 0.2),
    aes(x = farm, y = numbers, col = GrowthType)
             ) +
  scale_color_viridis_d() +
  scale_y_log10() +
  theme(strip.text.x = element_text(size = 8)) +
  ggtitle("Observed Bird Numbers", subtitle = "Colored by Vegetation Growth Type")
p + ggforce::facet_wrap_paginate(vars(species), nrow = 4, ncol = 13, page = 1)
p + ggforce::facet_wrap_paginate(vars(species), nrow = 4, ncol = 13, page = 2)
p + ggforce::facet_wrap_paginate(vars(species), nrow = 4, ncol = 13, page = 3)
```

```{r growthtypeonbirds_p, fig.width = 15, fig.height = 12}
obsb %>%
  pivot_longer(cols = names(bird_richness), names_to = "species", values_to = "detected") %>%
  filter(!is.na(GrowthType)) %>%
  group_by(species, GrowthType, farm, SurveyYear) %>%
  summarise(PropDetected = sum(detected) / n()) %>% #proportion within a single season at a farm
  summarise(PropDetected = mean(PropDetected), #across 'independent' seasons
            df = n()) %>%
  summarise(PropDetected = mean(PropDetected), #across 'independent' farms
            df = sum(df)) %>%
  mutate(se = sqrt(PropDetected * (1 - PropDetected) / df)) %>%
  inner_join(traits, by = c(species = "common_name")) %>%
  filter(!is.na(movements)) %>%
  ggplot() +
  geom_col(position = "dodge2",
           aes(x = species, y = PropDetected,
               group = GrowthType,
               fill = GrowthType)) +
  geom_errorbar(position = "dodge2",
           aes(x = species, ymin = PropDetected - 2*se,
               ymax = PropDetected + 2*se,
               group = GrowthType)) +
  scale_fill_viridis_d() +
  coord_flip() +
  facet_wrap(vars(movements), scales = "free_y") +
  theme(legend.position = "bottom") +
  ylab("Proportion of Visits at which Species Detected") +
  ggtitle("Birds: Detection")
```

There are many species for which growth type could be important: Red-Capped Robin, Brown Treecreeper...

GrowthType is plausibly available as maps across the full landscape. It is also possible that GrowthType is correlated with other predictors.

Error bars in above are *very rough*. They assume that revisits within a year at a single farm are highly correlated, and that vists on different seasons and different farms are independent. They are also based on a Gaussian approximation of 95% confidence intervals. 
__[I haven't checked my maths for this; very quick and dirty]__

```{r growthtypeonbirds_a, fig.width = 15, fig.height = 12}
obsa %>%
  pivot_longer(cols = names(bird_richness), names_to = "species", values_to = "numbers") %>%
  filter(!is.na(GrowthType)) %>%
  filter(numbers > 0) %>%
  inner_join(traits, by = c(species = "common_name")) %>%
  filter(movements == "Sedentary") %>%
  ggplot() +
  geom_violin(aes(GrowthType, numbers),
              draw_quantiles = c(0.25, 0.5, 0.75)) +
  # geom_point(
  #   position = position_jitter(width = 0.4, height = 0),
  #   aes(x = GrowthType, y = numbers, col = SurveyYear)) +
  # scale_color_viridis() +
  scale_y_log10() +
  coord_flip() +
  facet_wrap(vars(species)) +
  # theme(legend.position="bottom") +
  ggtitle("Sedentary Birds: Abundance vs GrowthType")
```


## Cropping and Grazing
```{r croppingvsgrazing, fig.width = 15, fig.height = 12}
sws_landuse <- read.table("./private/data/raw/LandUses SWS Farms.csv",
                          sep = ",", skip = 1,
                          as.is = TRUE)
colnames(sws_landuse) <- sws_landuse[1, ]
sws_landuse <- sws_landuse[-1, ]
sws_landuse$DmnntLandUse <- as.factor(sws_landuse$DmnntLandUse)

obsb %>%
  pivot_longer(cols = names(bird_richness), names_to = "species", values_to = "detected") %>%
  filter(!is.na(detected)) %>%
  inner_join(sws_landuse, by = c(farm = "FarmUnit")) %>%
  filter(!is.na(DmnntLandUse)) %>%
  group_by(species, DmnntLandUse, farm, SurveyYear) %>%
  summarise(PropDetected = sum(detected) / n()) %>% #proportion within a single season at a farm
  summarise(PropDetected = mean(PropDetected), #across 'independent' seasons
            df = n()) %>%
  summarise(PropDetected = mean(PropDetected), #across 'independent' farms
            df = sum(df)) %>%
  mutate(se = sqrt(PropDetected * (1 - PropDetected) / df)) %>%
  inner_join(traits, by = c(species = "common_name")) %>%
  filter(!is.na(movements)) %>%
  ggplot() +
  geom_col(position = "dodge2",
           aes(x = species, y = PropDetected,
               group = DmnntLandUse,
               fill = DmnntLandUse)) +
  geom_errorbar(position = "dodge2",
           aes(x = species, ymin = PropDetected - 2*se,
               ymax = PropDetected + 2*se,
               group = DmnntLandUse)) +
  scale_fill_viridis_d() +
  coord_flip() +
  facet_wrap(vars(movements), scales = "free_y") +
  theme(legend.position = "bottom") +
  ylab("Proportion of Visits at which Species Detected") +
  ggtitle("Birds: Detection")
```

There are multiple species where the farms labelled cropping, grazing or mixed have very different detection rates. 
Cropping vs Grazing vs Mixed may have statistically significant affect on: Noisy Miner, White-plumed Honeyeater, Dusky Woodswallow, Weebill, Jacky Winter, Grey Shrike-thrush, Galah...
Error bars in above are *very rough*. They assume that revisits within a year at a single farm are highly correlated, and that vists on different seasons and different farms are independent. They are also based on a Gaussian approximation of 95% confidence intervals. 
__[I haven't checked my maths for this; very quick and dirty]__