---
title: "Visualising SWS Bird Abundance Data"
author: "Kassel Hingee"
date: "14/02/2020"
output: 
  html_document: 
    fig_caption: yes
    toc: yes
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
out <- lapply(c("sf", "tsibble", 'lubridate', "viridis",
                'ggplot2', 'tidyr', 'grid', 'gridExtra', 
                'feasts', 'dplyr', 'gtable', 'fable',
                'mgcv'),
       library, character.only = TRUE)
```

## Import Data
```{r importdata}
bird_richness <- readRDS("./private/data/clean/sws_bird_richness.rds") # contains all bird spp
birds <- readRDS("./private/data/clean/sws_birds.rds") # contains only common bird spp
stopifnot(all.equal(bird_richness[, names(birds)], birds))

sites_rs <- readRDS("./private/data/clean/sws_sites_rs.rds") %>%
    mutate(farm = factor(substr(SiteCode, 1, 4)),
         sitenum = factor(as.integer(substr(SiteCode, 5, 5))))
sites_rs$log_plus_one_woody_cover <- log(sites_rs$woody_cover + 1)
traits <- readRDS("./private/data/clean/sws_traits.rds")

obsa <- cbind(sites_rs, bird_richness) 
```

```{r makepresenceabsence}
# make binary
for(i in 1:ncol(birds)){
  birds[which(birds[, i] > 0), i] <- 1
}
for(i in 1:ncol(bird_richness)){
  bird_richness[which(bird_richness[, i] > 0), i] <- 1
}
obsb <- cbind(sites_rs, bird_richness)
```

## Visualise Abundance
```{r vis_obsa_magpies}
obsa %>%
  #filter(farm == "BELL") %>%
  ggplot() +
  facet_wrap(vars(farm)) +
  geom_point(position = "jitter", #position_jitter(width = 1E6, height = 0),
    aes(x = SurveyDate, y = `Australian Magpie`)
             ) +
  scale_y_log10() +
  ggtitle("Australian Magpie Numbers Observed")
```

Numbers of detected Austaralian Magpie appears to have roughly constant density.

```{r vis_obsa_apostlebird}
obsa %>%
  #filter(farm == "BELL") %>%
  ggplot() +
  facet_wrap(vars(farm)) +
  geom_point(position = "jitter", #position_jitter(width = 1E6, height = 0),
    aes(x = SurveyDate, y = `Apostlebird`)
             ) +
  scale_y_log10() +
  ggtitle("Apostlebird Numbers Observed")
```


```{r vis_obsa_apostlebird_b}
obsa %>%
  #filter(farm == "BELL") %>%
  ggplot() +
  # facet_wrap(vars(farm)) +
  geom_point(position = "jitter", #position_jitter(width = 1E6, height = 0),
    aes(x = farm, y = `Apostlebird`, col = year(SurveyDate))
             ) +
  scale_color_viridis(name = "Survey Year") +
  scale_y_log10() +
  ggtitle("Apostlebird Numbers Observed")

obsa[, "SurveyDate", drop = FALSE]
```

That the same farms have a variety of colours suggests that the things that are changing in time are not an enormous factor on Apostlebirds.


```{r vis_obsa_all}
p <- obsa %>%
  pivot_longer(cols = names(bird_richness), names_to = "species", values_to = "numbers") %>%
  #filter(farm == "BELL") %>%
  ggplot() +
  # facet_wrap(vars(species)) +
  # ggforce::facet_wrap_paginate(vars(species), nrow = 4, ncol = 13) +
  geom_point(position = "jitter", #position_jitter(width = 1E6, height = 0),
    aes(x = farm, y = numbers, col = year(SurveyDate))
             ) +
  scale_color_viridis(name = "Survey Year") +
  scale_y_log10() +
  theme(strip.text.x = element_text(size = 8)) +
  ggtitle("Observed Bird Numbers")
p + ggforce::facet_wrap_paginate(vars(species), nrow = 4, ncol = 13, page = 1)
p + ggforce::facet_wrap_paginate(vars(species), nrow = 4, ncol = 13, page = 2)
p + ggforce::facet_wrap_paginate(vars(species), nrow = 4, ncol = 13, page = 3)
```


Looks like the Noisy Friarbird was widely around in early years and later years, but not years around 2010.
There are a number of species that were only observed much less in later years (some of this could be confounded with any farms/sites that dropped out of the study).
