---
title: "Compatability of Garnett et al's Australian Bird Data"
author: "Kassel Hingee"
date: "02/06/2020"
output:
  html_document: 
    collapsed: no
    df_print: kable
    number_sections: yes
    toc: yes
    toc_float: yes
---

Need to check:

+ names in Garnett's data match the names in our data
+ the removal set given by Garnett's data is consistent with the removal set given by Ikin's data

```{r prep, include = FALSE}
library(dplyr); library(tidyselect); library(tidyr);
knitr::opts_knit$set(root.dir = rprojroot::find_root(rprojroot::has_file("DESCRIPTION")))
devtools::load_all(rprojroot::find_root(rprojroot::has_file("DESCRIPTION")))
```

```{r importdata}
traits <- read.csv("./private/data/raw/Australian_Bird_Data_Version_1.csv")
traits_ikin <- as.data.frame(
  readxl::read_excel(
    "./private/data/raw/Ikin_SWS_Bird_Traits_updatedApril2017.xlsx",
    sheet = "Ikin_SWS_Bird_Traits"
  ))
birds_raw <- readRDS("./private/data/raw/birds_long.rds")
birds_raw$CommonName <- gsub("Horsfield\x92s Bushlark", "Horsfield's Bushlark", birds_raw$CommonName) #weird encoding of a bird name
birds_raw <- birds_raw %>% dplyr::filter(CommonName != "Corella sp.") #remove the single observation of Corella sp. as this could be at least two different common names.
```

## The names in Garnett's data match the names in our data
### Unique names for rows
Column's in Garnett's data that make for good names are unique to each taxon (each row), and similar common names or scientific names in our data.

```{r columnthataregoodnames}
anyDuplicated(traits$X3_Taxon_common_name_2)
anyDuplicated(traits$X5_Species_name_2)
anyDuplicated(traits[, c("X5_Species_name_2", "X6_Subspecies_name_2")])
anyDuplicated(traits$X10_Family_scientific_name_2)
anyDuplicated(traits$X17_Species_2)
anyDuplicated(traits[, c("X17_Species_2", "X18_Subspecies_2")])
```

In fact only the index and X3_Taxon_common_name_2 are unique to each row:
```{r colswithuniquevals}
which(apply(traits, 2, anyDuplicated) == 0)
```

`X3_Taxon_common_name_2` is also quite similar to our common names (see below).

### Mapping to our CommonName
Generalised Levenshtein distances between names
```{r namedists}
namedists <- adist(unique(birds_raw$CommonName), traits$X3_Taxon_common_name_2)
rownames(namedists) <- unique(birds_raw$CommonName)
colnames(namedists) <- traits$X3_Taxon_common_name_2
```

Summarising the row minimums shows that nearly every common name we have in our data is within (approximately) 1 substitution from a name in the Australia Birds data set
```{r sum_namedists}
summary(factor(apply(namedists, 1, min, na.rm = TRUE)))
```

Create a mapping by get the closest taxon names for each of our common names
```{r makeNameMap}
NameMap <- apply(namedists, 1, function(v){
  ind <- which.min(v)
  return(names(v)[[ind]])
})
```
names of NameMap are the common name that we use in our data
values of NameMap are the common names in the Garnett et al data set

These species have a difference of 1 `r NameMap[apply(namedists, 1, min) == 1]`.

These names have a higher difference `r NameMap[apply(namedists, 1, min) > 1]`.

Transform our data:

```{r bird_renamed}
birds_renamed <- birds_raw %>%
  mutate(CommonName = NameMap[CommonName])
```

### A quick note on Corella Sp
Corella sp has been removed because it could be multiple species
```{r corella1}
grep("Corella", unique(birds_renamed$CommonName), ignore.case = TRUE, value = TRUE)
grep("Corella", traits$X3_Taxon_common_name_2, ignore.case = TRUE, value = TRUE)
```

## Birds to Exclude (and comparison to using Ikin's)
Exclude birds taxonimically: raptors and obligate water birds.
We will also exclude the Australian Reed-Warbler, which is a passerine, because it requires wetlands to persist.

```{r birdsexlcude}
excluded_orders <- c(
  "Caprimulgiformes", # Frogmouths
  "Accipitriformes", # Eagles, Kites, Goshawks and Osprey
  "Strigiformes", # Masked Owls
  "Falconiformes", # Falcons
  "Podicepiformes", # Greebes
  "Gruiformes", # Crakes, Rails and Swamphens
  "Pelecaniformes", # Herons, Egrets, Pelicans etc
  "Anseriformes" # Ducks, Geese
)

birds_to_exclude <- as_tibble(traits) %>%
  dplyr::filter((X11_Order_2 %in% excluded_orders) | (X3_Taxon_common_name_2 == "Australian Reed-Warbler")) %>%
  dplyr::distinct()

# Birds to Include if Observed
birdskeep <- traits[!(traits$X1_Taxon_sort_1 %in% birds_to_exclude$X1_Taxon_sort_1), ]
```

To compare to Ikin need to create a map between species (taxon) names
```{r IkinNameMap}
namedists_ikin <- adist(traits_ikin$species, traits$X3_Taxon_common_name_2)
rownames(namedists_ikin) <- unique(traits_ikin$species)
colnames(namedists_ikin) <- traits$X3_Taxon_common_name_2
NameMap_ikin <- apply(namedists_ikin, 1, function(v){
  ind <- which.min(v)
  return(names(v)[[ind]])
})
# names of NameMap_ikin are the common name that Ikin's traits use
# values of NameMap_ikin are the common names in the Garnett et al data set
# The following have a difference of 1
NameMap_ikin[apply(namedists_ikin, 1, min) > 0]
```

Transform Ikin:
```{r transformIkinTraits}
traits_ikin$species <- NameMap_ikin[traits_ikin$species]
```

Compare these to Ikin
```{r Ikinremove}
species_to_remove_ikin <- traits_ikin %>%
  filter((diet == "Vertebrates") |
           (substrate == "Water") |
           (species == "Australian Reed-Warbler")) %>%
  dplyr::select(species) %>%
  arrange(species) %>%
  unlist() %>%
  as.vector()
species_to_keep_ikin <- setdiff(traits_ikin$species, species_to_remove_ikin)
```



### Compare species removed
nearly all birds we want to remove based on Ikin's traits are in the birds to exclude list based on Garnett's data
```{r vsIkinRemove}
summary((species_to_remove_ikin %in% birds_to_exclude$X3_Taxon_common_name_2))
species_to_remove_ikin[!(species_to_remove_ikin %in% birds_to_exclude$X3_Taxon_common_name_2)]
```

### Compare species not removed using Ikin
How about species suggested to keep using Ikin's traits, but excluded by the new method above?
```{r vsIkinKeep}
summary(species_to_keep_ikin %in% birds_to_exclude$X3_Taxon_common_name_2)
species_to_keep_ikin[(species_to_keep_ikin %in% birds_to_exclude$X3_Taxon_common_name_2)]
birds_renamed %>%
  dplyr::filter(CommonName %in% species_to_keep_ikin[(species_to_keep_ikin %in% birds_to_exclude$X3_Taxon_common_name_2)]) %>%
  group_by(CommonName) %>%
  summarise(Nobsatvisits = n()) %>%
  arrange(-Nobsatvisits)
```

11 species kept using Ikin's traits are removed using the new method. They look like ducks and water birds.
Only 6 of these are observed enough to plausibly be placed in our model.

The following compare these 11 species on some functional traits: they all feed on inland waters. Some eat vertebrates.

```{r ikinskeep_garnettexclude_summproperties}
library(ggplot2)
birds_to_exclude %>%
  dplyr::filter(X3_Taxon_common_name_2 %in% species_to_keep_ikin) %>%
  dplyr::select(X3_Taxon_common_name_2, X169_Food_Terrestrial_vertebrates_10,
                  X157_Breeding_habitat_inland_wetland_9,
                  matches("_Feeding_habitat_Inland_waters_")) %>%
  pivot_longer(-X3_Taxon_common_name_2) %>%
  dplyr::filter(value == 1) %>%
  ggplot() +
  geom_point(aes(y = X3_Taxon_common_name_2, x = name, col = name))
```


## 7_2_xx Modelled Species with Ikin Traits
```{r modelledspecies_ikintraits}
modeldata <- readRDS("~/dlinkingdatacode/private/data/clean/7_2_10_input_data.rds")
modelledspecies <- modeldata$species
length(setdiff(modelledspecies, traits_ikin$species))

traits_ikin[traits_ikin$species %in% modelledspecies, ]
```

All species modelled have traits in the traits_ikin data! That is very convenient for functional diversity!

## Compare to DBO Traits
```{r load_dbo_traits}
source('../../../ComputerRelated/AccessLindenmayerSQL.R', echo=FALSE)

dbo_traits <- dbGetQuery(con, 
"SELECT sp.CommonName, bs.*
FROM dbo.BirdSpecies AS bs
INNER JOIN dbo.Species AS sp
ON bs.SpeciesId=sp.SpeciesId")
DBI::dbDisconnect(con)
```

```{r getscoresformodelledbirds}
all(modelledspecies %in% dbo_traits$CommonName)
dbo_traits %>%
  dplyr::filter(CommonName %in% modelledspecies) %>%
  dplyr::select(c(CommonName, ends_with("Score"))) %>%
  filter_all(any_vars(is.na(.)))
```

There are five birds that don't have any trait scores information:
White-throated Gerygone				
Yellow-faced Honeyeater				
White-plumed Honeyeater				
White-plumed Honeyeater				
Diamond Firetail		

```{r checkoutmissing_in_full}
dbo_traits %>%
  dplyr::filter(CommonName %in% c(
    "White-throated Gerygone",			
    "Yellow-faced Honeyeater",			
    "White-plumed Honeyeater",			
    "White-plumed Honeyeater",			
    "Diamond Firetail"
  )) %>%
  dplyr::filter(PresentInRegion == "1") %>%
  dplyr::select(c(CommonName, ends_with("Score")))
```

Some CommonNames occur twice or three times! Many of species present in region versions have scores.
Of the 5 common names, only Yellow-faced Honeyeater has no scores.
