---
title: "Compatability of Garnett et al's Australian Bird Data"
author: "Kassel Hingee"
date: "02/06/2020"
output: 
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
---

Need to check:

+ names in Garnett's data match the names in our data
+ the removal set given by Garnett's data is consistent with the removal set given by Ikin's data

```{r prep}
library(dplyr); library(tidyselect); library(tidyr);
```

```{r importdata}
traits <- read.csv("./private/data/raw/Australian_Bird_Data_Version_1.csv")
traits_ikin <- as.data.frame(
  readxl::read_excel(
    "./private/data/raw/Ikin_SWS_Bird_Traits_updatedApril2017.xlsx",
    sheet = "Ikin_SWS_Bird_Traits"
  ))
birds_raw <- readRDS("./private/data/raw/birds_long.rds")
birds_raw$CommonName <- gsub("Horsfield\x92s Bushlark", "Horsfield's Bushlark", birds_raw$CommonName) #weird encoding of a bird name
birds_raw <- birds_raw %>% dplyr::filter(CommonName != "Corella sp.") #remove the single observation of Corella sp. as this could be at least two different common names.
```

## The names in Garnett's data match the names in our data
### Unique names for rows
Column's in Garnett's data that make for good names are unique to each taxon (each row), and similar common names or scientific names in our data.

```{r columnthataregoodnames}
anyDuplicated(traits$X3_Taxon_common_name_2)
anyDuplicated(traits$X5_Species_name_2)
anyDuplicated(traits[, c("X5_Species_name_2", "X6_Subspecies_name_2")])
anyDuplicated(traits$X10_Family_scientific_name_2)
anyDuplicated(traits$X17_Species_2)
anyDuplicated(traits[, c("X17_Species_2", "X18_Subspecies_2")])
```

In fact only the index and X3_Taxon_common_name_2 are unique to each row:
```{r colswithuniquevals}
which(apply(traits, 2, anyDuplicated) == 0)
```

`X3_Taxon_common_name_2` is also quite similar to our common names (see below).

### Mapping to our CommonName
Generalised Levenshtein distances between names
```{r namedists}
namedists <- adist(unique(birds_raw$CommonName), traits$X3_Taxon_common_name_2)
rownames(namedists) <- unique(birds_raw$CommonName)
colnames(namedists) <- traits$X3_Taxon_common_name_2
```

Summarising the row minimums shows that nearly every common name we have in our data is within (approximately) 1 substitution from a name in the Australia Birds data set
```{r sum_namedists}
summary(factor(apply(namedists, 1, min, na.rm = TRUE)))
```

Create a mapping by get the closest taxon names for each of our common names
```{r makeNameMap}
NameMap <- apply(namedists, 1, function(v){
  ind <- which.min(v)
  return(names(v)[[ind]])
})
```
names of NameMap are the common name that we use in our data
values of NameMap are the common names in the Garnett et al data set

These species have a difference of 1 `r NameMap[apply(namedists, 1, min) == 1]`.

These names have a higher difference `r NameMap[apply(namedists, 1, min) > 1]`.

### A quick note on Corella Sp
Corella sp has been removed because it could be multiple species
```{r corella1}
grep("Corella", unique(birds_raw$CommonName), ignore.case = TRUE, value = TRUE)
grep("Corella", traits$X3_Taxon_common_name_2, ignore.case = TRUE, value = TRUE)
```

## vs Birds to Exclude using Ikin's
feed in water;
breed in wetlands;
birds that eat vertebrates;
birds that nest on floating platforms

```{r birdsexlcude}
waterfeeders <- as_tibble(traits) %>%
  dplyr::filter_at(vars(matches("_Feeding_habitat_Inland_waters_")), any_vars(. == 1))
wetlandbreeders <- as_tibble(traits) %>%
  dplyr::filter(X157_Breeding_habitat_inland_wetland_9 == 1)

vertebrateeaters <- as_tibble(traits) %>%
  dplyr::filter(X169_Food_Terrestrial_vertebrates_10 == 1) #warning lots of NA values (50%)

floatingnesters <- as_tibble(traits) %>%
  dplyr::filter(X188_Nest_location_Floating_12 == 1)

birds_to_exclude <- bind_rows(waterfeeders,
                              wetlandbreeders,
                              vertebrateeaters,
                              floatingnesters) %>%
  dplyr::distinct()

# Birds to Include if Observed
birdskeep <- traits[!(traits$X1_Taxon_sort_1 %in% birds_to_exclude$X1_Taxon_sort_1), ]
```

Compare these to Ikin 
```{r Ikinremove}
species_to_remove_ikin <- traits_ikin %>%
  filter((diet == "Vertebrates") |
           (substrate == "Water") |
           (species == "Australian Reed-Warbler")) %>%
  dplyr::select(species) %>%
  arrange(species) %>%
  unlist() %>%
  as.vector()
species_to_keep_ikin <- setdiff(traits_ikin$species, species_to_remove_ikin)
```

To compare need to create a map between species (taxon) names
```{r IkinNameMap}
namedists_ikin <- adist(traits_ikin$species, traits$X3_Taxon_common_name_2)
rownames(namedists_ikin) <- unique(traits_ikin$species)
colnames(namedists_ikin) <- traits$X3_Taxon_common_name_2
NameMap_ikin <- apply(namedists_ikin, 1, function(v){
  ind <- which.min(v)
  return(names(v)[[ind]])
})
# names of NameMap_ikin are the common name that Ikin's traits use
# values of NameMap_ikin are the common names in the Garnett et al data set
# The following have a difference of 1
NameMap_ikin[apply(namedists_ikin, 1, min) > 0]
```

### Compare species removed
nearly all birds we want to remove based on Ikin's traits are in the birds to exclude list based on Garnett's data
```{r vsIkinRemove}
summary((NameMap_ikin[species_to_remove_ikin] %in% birds_to_exclude$X3_Taxon_common_name_2))
species_to_remove_ikin[!(NameMap_ikin[species_to_remove_ikin] %in% birds_to_exclude$X3_Taxon_common_name_2)]
```

The exception, the Eastern Great Egret, has no data on any feeding habitat, breeding habitat or food. Ikin regards it as having a substrate of water.
```{r easterngreategret}
traits[traits$X3_Taxon_common_name_2 == "Eastern Great Egret", ] %>%
  dplyr::select(matches("(_9|_10)$")) %>%
  t()
traits[traits$X3_Taxon_common_name_2 == "Eastern Great Egret", ] %>% t()
traits_ikin[traits_ikin$species == "Eastern Great Egret", ]
```

It doesn't matter as the Eastern Great Egret is too rare to include in our current method
```{r easterngreategret_observed}
birds_raw[birds_raw$CommonName == "Eastern Great Egret", ]
```

### Compare species not removed using Ikin
How about species suggested to keep using Ikin's traits?
```{r vsIkinKeep}
summary((NameMap_ikin[species_to_keep_ikin] %in% birds_to_exclude$X3_Taxon_common_name_2))
species_to_keep_ikin[(NameMap_ikin[species_to_keep_ikin] %in% birds_to_exclude$X3_Taxon_common_name_2)]
```

This is a bit troubling. My filtering using Garnett's traits means removing magpies, kookaburras and more (38 in total).

```{r ikinskeep_garnettexclude_summproperties}
library(ggplot2)
birds_to_exclude %>%
  dplyr::filter(X3_Taxon_common_name_2 %in% NameMap_ikin[species_to_keep_ikin]) %>%
  dplyr::select(X3_Taxon_common_name_2, X169_Food_Terrestrial_vertebrates_10,
                  X157_Breeding_habitat_inland_wetland_9,
                  matches("_Feeding_habitat_Inland_waters_")) %>%
  pivot_longer(-X3_Taxon_common_name_2) %>%
  dplyr::filter(value == 1) %>%
  ggplot() +
  geom_point(aes(y = X3_Taxon_common_name_2, x = name, col = name))
```

Many of the inconsistent birds eat terrestrial vertebrates. 

```{r ikinskeep_garnettexclude}
# many of the 38 inconsistent birds also eat vertebrates, about a third or all excluded birds eat vertebrates
birds_to_exclude %>% select(X169_Food_Terrestrial_vertebrates_10) %>% sum()
birds_to_exclude %>%
  dplyr::filter(X3_Taxon_common_name_2 %in% NameMap_ikin[species_to_keep_ikin]) %>%
  dplyr::select(matches("_10$")) %>%
  dplyr::filter(X169_Food_Terrestrial_vertebrates_10 == 1) #warning lots of NA values (50%)

# half the excluded birds feed in agricultural landscapes, yet 97% of inconsistent birds use agrigulutral landscapes to feed
birds_to_exclude %>% select("X144_Feeding_habitat_Agricultural_landscapes_9") %>% sum()
birds_to_exclude %>% 
  dplyr::filter(X3_Taxon_common_name_2 %in% NameMap_ikin[species_to_keep_ikin]) %>%
  select("X144_Feeding_habitat_Agricultural_landscapes_9") %>% colMeans()
```

