---
title: "Check Extracted Remote Sensing Data"
author: "Kassel Hingee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    collapsed: no
    number_sections: yes
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_root(rprojroot::has_file("DESCRIPTION")))
devtools::load_all(rprojroot::find_root(rprojroot::has_file("DESCRIPTION")))
library(sf)
library(leaflet)
library(ggplot2)
```

```{r importdata}
data_7_2_7 <- readRDS("./private/data/clean/7_2_4_input_data.rds")
Xocc <- rbind(data_7_2_7$insampledata$Xocc, data_7_2_7$holdoutdata$Xocc)
yXobs_grnd <- rbind(data_7_2_7$insampledata$yXobs, data_7_2_7$holdoutdata$yXobs)
species <- data_7_2_7$species

locs <- read.csv("private/data/clean/site_locations_cleaned.csv", stringsAsFactors = FALSE)
locs <- st_as_sf(locs, coords = c("MeanLON", "MeanLAT"))
locs <- st_set_crs(locs, 4326) #4326 is the epsg code for WGS84, make this default crs I'll return to

gpp <- readRDS("./private/data/remote_sensed/8d_gpp.rds")
woody <- readRDS("./private/data/remote_sensed/woody500.rds")
twi <- readRDS("./private/data/remote_sensed/TWI_boxgum_sites.rds")
```

Components to check:

+ Mean GPP for each location
+ TWI for each location
+ Woody veg per year (i.e. visit)
+ GPP difference for each visit


## TWI for each location
```{r twi_hist}
twi %>%
  ggplot() +
  geom_histogram(aes(x = TWI), bins = 40)
twi %>%
  filter(TWI > 12)
```

There are 6 sites with extremely high TWI values.

```{r TWI1}
twi_loc <- st_as_sf(inner_join(twi, locs[, "SiteCode"], by = "SiteCode"))

pal <- colorNumeric(viridisLite::viridis(10), twi_loc$TWI)
map <- leaflet() %>%
  # addTiles() %>%
  # addProviderTiles("Esri.WorldImagery") %>%
  # addProviderTiles("Esri.DeLorme") %>%
  addProviderTiles("Esri.WorldShadedRelief") %>%
  addScaleBar(options = scaleBarOptions(maxWidth = 200, imperial = FALSE)) 
  # addOpacitySlider("elevation")
map <- map %>%
  addCircleMarkers(lng = st_coordinates(twi_loc)[, "X"], lat = st_coordinates(twi_loc)[, "Y"],
                   label = paste(twi_loc$TWI, twi_loc$SiteCode),
                   color = pal(twi_loc$TWI),
                   opacity = 1) %>%
  addLegend(pal = pal, values = seq(min(twi_loc$TWI), max(twi_loc$TWI), length = 20),
            title = "TWI")
map
```

The TWI at CLOT-S is unusally high. Neighbouring CLOT-C is much lower and there seems to be little elevation difference between the two.
It appears, roughly, that nearby points higher on the Earth's surface have lower TWI.

Check against a raster map.

```{r TWI2}
library(slga) #specialist library for reading the data
library(raster)
library(rasterVis)
library(leaflet.opacity)
twi_ras <- get_lscape_data(slga_product_info %>%
                   dplyr::filter(Product == "Topographic Wetness Index") %>%
                   dplyr::select(Short_Name) %>%
                   as.character(),
                   aoi = locs[locs$StudyCode == "Nanangroe Natural Experiment", ])

map %>%
  fitBounds(extent(twi_ras)@xmin, extent(twi_ras)@ymin, extent(twi_ras)@xmax, extent(twi_ras)@ymax) %>%
  addRasterImage(twi_ras, colors = pal, layerId = "raster") %>%
  addOpacitySlider(layerId = "raster")

# gplot(twi_ras) + 
#   geom_tile(aes(fill = value)) +
#   scale_fill_viridis_c() +
#   geom_sf(data = locs[locs$StudyCode == "Nanangroe Natural Experiment", ], inherit.aes = FALSE, stat = "sf") +
#   coord_sf()
# ggplot() + geom_sf(data = twi_loc)
```

A raster extraction of TWI has values consistent with the values at the site locations.
The raster map is also consistent with the ESRI terrain, confirming that the projections are approximately correct.

## Woody Vegetation, 500m Buffer
```{r woodyveghist}
anyNA(woody, recursive = TRUE) #there are no NA values in the woody data

woody %>% 
  pivot_longer(-SiteCode, names_to = "year", values_to = "woody500m") %>%
  filter(iis(woody500m))

woody %>% 
  pivot_longer(-SiteCode, names_to = "year", values_to = "woody500m") %>%
  ggplot() +
  geom_histogram(aes(x = woody500m), bins = 50)

woody %>% 
  pivot_longer(-SiteCode, names_to = "year", values_to = "woody500m") %>%
  ggplot() +
  facet_wrap(~year, scales = "free_y") +
  geom_histogram(aes(x = woody500m), bins = 30)

woody %>% 
  pivot_longer(-SiteCode, names_to = "year", values_to = "woody500m") %>%
  dplyr::filter(woody500m <= 1) #less than 1% 60ish site years
```

There are some very low (<1%) woody cover locations and years.
There are also some very high (>60%) woody cover data points.
Some of these anomalies may not be present in the data we are interested in because we are not interested in every year at every location.

The standard deviation of the woody veg values for each location will give an idea of any anomolous locations that have clearing or planting nearby.

```{r sdofwoodyveg}
woody %>% 
  pivot_longer(-SiteCode, names_to = "year", values_to = "woody500m") %>%
  group_by(SiteCode) %>%
  summarise(sd = sd(woody500m)) %>%
  ggplot() +
  geom_point(aes(x = sd, y = SiteCode))

woody %>% 
  pivot_longer(-SiteCode, names_to = "year", values_to = "woody500m") %>%
  group_by(SiteCode) %>%
  summarise(sd = sd(woody500m)) %>%
  filter(sd > 7)

woody %>% 
  pivot_longer(-SiteCode, names_to = "year", values_to = "woody500m") %>%
  filter(SiteCode %in% c("WEBB-2", "KEA-1", "WISE-B")) %>%
  ggplot() +
  facet_wrap(vars(SiteCode)) +
  geom_point(aes(x = year, y = woody500m))
```

The site WISE-B has an incredibly large sample standard deviation. This is driven by low woody veg values from 2003 through to 2010, and high woody veg values in other years. WISE-B is at the edge of a lake-like section of the Murray river, near Albury, could it be that the values of woody veg are not filtering out water well?

WEBB-2, also an outlier, appears to neighbour a plantation forest.

```{r WISE-B}
tile <- tilearoundobj(locs[locs$SiteCode == "WISE-B", ], 0.01)
wiseb_region <- brick_woodycover(as_Spatial(tile), 2000:2019)
wiseb_region[wiseb_region > 100] <- 0 #assume NA values mean no woody veg
plot(wiseb_region)

wf <- raster::focalWeight(wiseb_region, d = 500, type = "circle")
woody_bs <- focal_bylayer(wiseb_region, wf, fun = sum)
names(woody_bs) <- names(wiseb_region)
WISEB_woody_vals <- extract(woody_bs, as_Spatial(locs[locs$SiteCode == "WISE-B", ]), df = TRUE)
```

WISE-B's woody veg values are heavily affected by the nearby water body. The cover value appears to 157, and must represent NA. Replacing all values above 100 with NA will hopefully lead to a better result.

```{r histofwoodyatvisits}
names(yXobs_grnd)
```