---
title: "Inspecting Albers Tiles (for Woody Cover)"
author: "Kassel Hingee"
date: "09/12/2019"
output: html_document
---

Woody Cover is in Albers Tile format.
I think the coordinate scheme of these files is epsg:3577 (more later in this document on this).
The raster data is stored as tiles in folders xmin_ymin, where xmin and ymin are in '000 000s of meters Eastings and Northings.

This document is to check the settings required for the package *raster* to correctly read the tiles.
It tests chosen projection information using the very visually distinct location of Sydney.

It applies to an unofficially modified version of *raster* that contains partial bug fixes.
To install this version of raster run: remotes::install_github('https://github.com/kasselhingee/raster', ref = 'ce63b218')")
These fixes were not merged into the main raster package because they caused errors in other situations.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, root.dir = "..")
invisible(lapply(c("raster", "maptools", "rgdal", "ncdf4", "lubridate"),
                 library, character.only = TRUE))
invisible(lapply(paste0("./functions/", list.files("./functions")), source))
if (packageVersion("raster") != "3.0-7") {
    stop(paste("This demonstration uses the 'dims' argument of raster(). This argument requires an unofficial version of the raster package to work properly.",
    "To install this version of raster run:\n remotes::install_github('https://github.com/kasselhingee/raster', ref = 'ce63b218')"))
  }
```

# An arbitrarily chosen tile
```{r arbitrarytile}
nc <- nc_open("[fillmismatch]http://dapds00.nci.org.au/thredds/dodsC/ub8/au/LandCover/DEA_ALC/0_-12/fc_metrics_0_-12_2000.nc")
ncatt_get(nc, "crs")$crs_wkt
```

This text string cleaned up is:

        spatial_ref=PROJCS[
        "GDA94 / Australian Albers",
        GEOGCS["GDA94",DATUM["Geocentric_Datum_of_Australia_1994",SPHEROID["GRS 1980",6378137,298.257222101,AUTHORITY["EPSG","7019"]],TOWGS84[0,0,0,0,0,0,0],AUTHORITY["EPSG","6283"]],PRIMEM["Greenwich",0,AUTHORITY["EPSG","8901"]],UNIT["degree",0.0174532925199433,AUTHORITY["EPSG","9122"]],AUTHORITY["EPSG","4283"]],
        PROJECTION["Albers_Conic_Equal_Area"],
        PARAMETER["standard_parallel_1",-18],
        PARAMETER["standard_parallel_2",-36],
        PARAMETER["latitude_of_center",0]
        ,PARAMETER["longitude_of_center",132],
        PARAMETER["false_easting",0],
        PARAMETER["false_northing",0],
        UNIT["metre",1,AUTHORITY["EPSG","9001"]],
        AXIS["Easting",EAST],
        AXIS["Northing",NORTH],
        AUTHORITY["EPSG","3577"]]

This suggests projection coordinate system is called GDA94 this corresponds to epsg:3577 according to [https://spatialreference.org/ref/epsg/3577/]  (which is also given in last property in above list).
The datum (spheriod) is GRS1980.
The above fits with the projection type (Albers equal area) and the ellipsoidal argument in the proj.4 string
```{r GDA94}
CRS("+init=epsg:3577")
```

However EPSG:3577 has some differences to the CRS that *raster* extracts from the ncdf files
```{r bcrs}
r <- raster("[fillmismatch]http://dapds00.nci.org.au/thredds/dodsC/ub8/au/LandCover/DEA_ALC/0_-12/fc_metrics_0_-12_2000.nc", varname = "WCF")
proj4string(r)
```
The values for `proj`, `lat_1`, `lon_0`, `lat_0`, `x_0`, `y_0` are the same.
But the value of `lat_2` is different (45.5 vs -36) and the values
`a` and `rf` do not even appear in `CRS("+init=epsg:3577")`.


## Sydney
The longitude and latitude of Sydney is:
```{r sydlonglat}
syd_latlong <- SpatialPoints(matrix(c(151.209900,  -33.865143), nrow = 1 ), proj4string = CRS("+proj=longlat +datum=WGS84"))
coordinates(syd_latlong)
```

Transformed into epsg:3577:
```{r syd3577}
syd_epsg3577 <- spTransform(syd_latlong,
                            CRS("+init=epsg:3577"))
coordinates(syd_epsg3577)
```
This suggests tile 17_-39. The following reads in this data to check it

```{r sydtile}
bsyd <- raster("[fillmismatch]http://dapds00.nci.org.au/thredds/dodsC/ub8/au/LandCover/DEA_ALC/17_-39/fc_metrics_17_-39_2000.nc", varname = "WCF", dims = 2:1)
plot(bsyd)
plot(add = TRUE, syd_epsg3577)

sydzoom <- crop(bsyd, extent(buffer(syd_epsg3577, 5000)))
plot(sydzoom)
plot(add = TRUE, syd_epsg3577)
```

The above tile and point location match the location of Sydney very well.

Correcting the CRS of the raster object works:
```{r fixrastercrs}
proj4string(bsyd) <- CRS("+init=epsg:3577")
sydzoom <- crop(bsyd, extent(buffer(syd_epsg3577, 5000)))
plot(sydzoom)
plot(add = TRUE, syd_epsg3577)
```

### In projection as interpreted by *raster*:
This brings up errors quite quickly. Transformed Sydney location into the CRS interpreted by *raster* is
```{r sydraster}
syd_rasterCRS <- spTransform(syd_latlong,
            CRS("+proj=aea +lat_1=-18 +lon_0=132 +lat_0=0 +x_0=0 +y_0=0 +a=6378137 +rf=298.257222101 +lat_2=45.5"))
coordinates(syd_rasterCRS)
```
This means xmin_ymin is 21_-37, which doesn't even exist in the catalogue.


## Tiles Relevant to South-West Slopes

# Construct Region Desired
Our region of interest is:
```{r preproi}
sws_sites <- readRDS("./private/data/clean/sws_sites.rds")
points <- spTransform(sws_sites_2_spdf(sws_sites),
                       CRS("+init=epsg:3577"))
roi <- extent(points)
```

Xmins and ymins for a square covering the points:
```{r xmins}
tilestep <- 100000
lxmin <- floor(roi@xmin / tilestep) * tilestep #lowest xmin
xmins <- seq(lxmin, -1 + ceiling(roi@xmax / tilestep) * tilestep,
    by = tilestep)
lymin <- floor(roi@ymin / tilestep) * tilestep #lowest ymin
ymins <- seq(lymin, -1 + ceiling(roi@ymax / tilestep) * tilestep,
             by = tilestep)
```

Tiles are then
```{r tiles}
xmin_v_ymin <- expand.grid(xmin = xmins, ymin = ymins)
tilesidx <- apply(xmin_v_ymin / tilestep, 1, function(x) paste(x, collapse = "_"))
```

Files for the year 2000 are then:
```{r tilefiles_2000}
tilefiles2000 <- paste0("[fillmismatch]http://dapds00.nci.org.au/thredds/dodsC/ub8/au/LandCover/DEA_ALC/",
      tilesidx,
      "/fc_metrics_",
      tilesidx,
      "_2000.nc")
```

These rasters are:
```{r tileras}
r.l <- lapply(tilefiles2000,
              function(x){
                      tryCatch(
                        {ras <- raster(x, varname = "WCF", dims = 2:1)
                                     return(ras)
                        }
                      ,
                      warning = function(w) {
                              if (!grep("cannot process these parts of the CRS", as.character(w))){
                                  warning(paste("For", x, w))
                              } else {
                                  suppressWarnings(ras <- raster(x, varname = "WCF", dims = 2:1))
                              }
                      })
                      })
rall <- do.call(merge, r.l)
proj4string(rall) <- CRS("+init=epsg:3577")
plot(rall, maxpixels = 10000)
plot(add = TRUE, roi)
plot(add = TRUE, points)
```