---
title: "Investigating Log-Linear Model Fit"
author: "Kassel Hingee"
date: "10/01/2020"
output: 
  html_document: 
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, root.dir = "..")
out <- lapply(c("sf", "tsibble", 'lubridate', "viridis",
                'ggplot2', 'tidyr', 'grid', 'gridExtra', 
                'feasts', 'dplyr', 'gtable', 'fable'),
       library, character.only = TRUE)
out <- lapply(paste0("../functions/", list.files("../functions/")), source)
```

```{r preparedata, echo = FALSE}
#load data and convert to tsibbles
load("../private/data/remote_sensed/pg_daily.Rdata")
pg_daily$times <- as_date(pg_daily$times)
pg <- pg_daily %>%
  pivot_longer(-times, names_to = "site", values_to = "pg") %>%
  as_tsibble(key = site, index = times)
load("../private/data/remote_sensed/gpp_8d.Rdata")
gpp <- gpp_8d %>%
  pivot_longer(-times, names_to = "site", values_to = "gpp") %>%
  as_tsibble(key = site, index = times)
pggpp <- as_tsibble(dplyr::full_join(pg, gpp, by = c("times", "site")), key = site, index = times)
pggpp <- pggpp %>%
  mutate(yday = yday(times)) %>%
  group_by_key() %>% #key is site
  mutate(lininterp_gpp = zoo::na.approx(gpp, rule = 1, na.rm = FALSE))

pggpp_train <- pggpp %>%
  filter_index(. ~ "2016-12-31")
pggpp_test <- pggpp %>%
  filter_index("2017-01-01" ~ .)


pggpp_train_concatsites <- pggpp_train %>%
 mutate(siteindx = as.numeric(factor(site, levels = unique(pggpp$site)))) %>%
 mutate(alttimes = times - days(366 * 30 * siteindx)) %>% #can't use year periods due to leap years
 ungroup() %>%
 update_tsibble(key = NULL, index = "alttimes") %>%
 fill_gaps()
```

```{r loadfittedmodel, echo = TRUE}
fit5 <- readRDS("../analysis/fit_fmla5_concatsites.rds")
```

The model investigated here is a linear model of log(GPP + 0.001) with predictors:
```{r modelfmla, echo = FALSE}
tidy(fit5)$term
```

Note that that 0.001 is designed to avoid taking the logarithm of 0 (about 0.5% of all GPP measurements are 0). The value was chosen to be a tiny fraction of the range of observed GPP, which was from 0 to about 15.

The first terms are logarithms of linearly interpolated GPP (+  0.001) on the same day of the year for 1, 2, 3, 4, and 5 years in the past.
The intercept term is the usual to-be-fitted constant offset independent of predictors.
The other terms are the sum of precipitation across 7 day windows, lagged by 0, 7, 14, ..., 63 days (0 weeks through to 9 weeks).

In mathematical terminology the model is:
\[
\begin{aligned}
 \log(Y_i(t) + 0.001) & = & \theta_{10} + \theta_{11} \log(Y_i(t - 1yr) + 0.001) +  \theta_{12} \log(Y_i(t - 2yr) + 0.001) + \theta_{13}\log(Y_i(t - 3yr) + 0.001) +\\
&& \theta_{14}\log(Y_i(t - 4yr) + 0.001) + \theta_{15}\log(Y_i(t - 5yr) + 0.001)  +\\
&& \theta_{20} X_i(t) +  \theta_{22}X_i(t - 1w) + ... + \theta_{29}X_i(t - 9w) +
\epsilon_i(t)\\
\iff && \\
Y_i(t) + 0.001 & = & \exp(\theta_{10}) 
  (Y_i(t - 1yr) + 0.001)^{\theta_{11}}
  (Y_i(t - 2yr) + 0.001)^{\theta_{12}}
  (Y_i(t - 3yr) + 0.001)^{\theta_{13}}\\
&&  (Y_i(t - 4yr) + 0.001)^{\theta_{14}}
  (Y_i(t - 5yr) + 0.001)^{\theta_{15}} \\
&& \exp(X_i(t))^{\theta_{20}} 
  \exp(X_i(t - 1w))^{\theta_{21}}...
  \exp(X_i(t - 9w))^{\theta_{29}}
  \exp(\epsilon_i(t))\\
\end{aligned}
\]
where 

+ $Y_i$ is the gpp at site $i$
+ $X_i$ is the sum of rain in the last 7 days at site $i$
+ $t$ is the date
+ $\epsilon_i(t)$ is a collection of i.i.d random variables, each with Gaussian distribution with mean of $0$ and standard deviation independent of the site $i$.

This model can be viewed as a familiar log-linear model (the error term is assumed to have no autocorrelation). It was fitted using the `fable` function `ARIMA` which uses base R's `arima` function. However, I am fairly sure that fitting using `stats::lm()` would have worked also, but involved further coding.

The terms are based on the assumption that GPP is seasonal (i.e. can be predicted from the GPP on the same day in previous years) and that it is dependent on precipitation in the last 9 weeks.


## Fitted Parameters
The fitted parameters for this model are
```{r fittedparams, echo = FALSE}
print(tidy(fit5) %>% select(- .model), n = 17)

fit5 %>%
  tidy() %>%
  mutate(term = factor(term, levels = tidy(fit5)$term, ordered = TRUE)) %>%  #makes sure plot maintains order of terms
  filter(term != "Intercept") %>%
  ggplot() +
  #geom_bar(aes(x = term, y = estimate), stat = "identity") +
  geom_pointrange(aes(x = term, ymin = estimate - 2 * std.error, ymax = estimate + 2 * std.error, y = estimate),
                  stat = "identity", fatten = 0.1, na.rm = TRUE) +
  ylim(c(-0.1, 0.5)) +
  coord_flip()
```

Intercept is not shown here for convenience as the estimated intercept is much larger than the other coefficients.


*If* the model assumptions are correct then all the predictors are statistically significant (the above plot has the standard error range represented by segments).

## Residual analysis

```{r residuals_training, echo = FALSE}
est_resid_training <- inner_join(residuals(fit5), fitted(fit5))
stopifnot(all.equal(est_resid_training$alttimes, pggpp_train_concatsites$alttimes)) #checking that regenerated 
est_resid_training <- right_join(est_resid_training, pggpp_train_concatsites) %>%
                filter(!is.na(times)) %>%
                update_tsibble(key = site, index = times)


est_resid_training %>%
  ggplot() +
  geom_histogram(aes(x = .resid), na.rm = TRUE, bins = 80) +
  ggtitle("Histogram of Residuals: all days and sites combined")

est_resid_training %>%
  ggplot() +
  geom_abline(aes(intercept = 0, slope = 1), col = "grey", lwd = 0.5) +
  geom_qq(aes(sample = .resid), na.rm = TRUE) +
  ggtitle("Q-Q plot of residuals: all days and sites combined")

est_resid_training %>%
  mutate(lagloggpp = lag(log(lininterp_gpp + 0.001)), n = 365.25) %>%
  ggplot() +
  geom_bin2d(aes(x = lagloggpp, y = .resid), bins = 60) +
  scale_fill_viridis(trans = "log") +
  ggtitle("Density of residual occurence against GPP of one year earlier")
```

The histogram reveals a cluster of highly negative residuals.
The Q-Q plot is highly non-linear, confirming that residuals are not normally distributed.
From the plot of density of residual occurences against lagged log GPP:

a. the highly negative residuals occur only when the previous year has nearly 0 GPP (0.28 or smaller).
b. the highly negative residuals appear to be separate from the residuals (there are very few residuals bettween - 5 and -2.5, and these do not occur at the same lag log gpp).
c. the remaining residuals have a mean that depends on the lagged log GPP values; it is difficult to tell if the the variance of the residuals also depends on the value of lagged log GPP.

When the previous year has positive GPP, and the current year has 0 GPP. The *only* multiplicative models that can predict this involve multiplying by something close to $0 = \exp(-\infty)$. Thus we should expect any multiplicative models to have large residuals for these situations, noting that the residual is only large in *transformed* units, the residuals in natural GPP units is less than 0.5.

1. try adding larger numbers than 0.001 to reduce the emphasis (in terms of model fitting objective) on close-to-zero values of GPP. This feel's like a poor option though as it brings lots of questions about why the model has that form. It might be better to instead assume that the error terms $\epsilon_i$ has an extreme-value component and then use extreme-value methods (somehow).
2. OR exclude GPP observations lower than $exp(-2.5) - 0.001$ from the training data, and have a caveat that lower GPP values are not modelled well.


## Plots of forecasted values

### For the training data
```{r plotforecasted_training, echo = FALSE}
est_resid_training %>%
  filter(site %in% unique(pggpp$site)[c(1, 5, 9, 19, 54)]) %>%
  filter_index("2005" ~ .) %>%
  pivot_longer(c(.fitted, gpp), names_to = "valname") %>%
  as_tsibble(key = c(site,valname), index = times) %>%
  mutate(value = zoo::na.approx(value, rule = 1, na.rm = FALSE)) %>%
  ggplot() +
  facet_grid(rows = vars(site)) +
  geom_line(aes(x = times, y = value, col = valname, lty = valname), na.rm = TRUE) +
  ggtitle("Forecasts and Observed GPP: training data")
```

The fitted (forecast) values here have been automatically transformed back to GPP (technically the model fitted log(GPP + 0.001)). 
[evidence: the histogram of fitted values is similar to the histogram of observed GPP, and that the inverse transorfmation, $\exp(x) - 0.001$, of the fitted values is completely different to the historgram of observed GPP values]


The 2007 spring peak is significantly underestimated. Is this because GPP of previous year was so low and precipitation doesn't have a big impact in the model?

The first GPP peak in 2010 at GEDD2 is not forecasted (it also wasnt forecasted in the additive model). The rain causing this peak was only a few weeks earlier, suggesting that it something about the way precipitation is incorporated into the model.

1. correlation analysis, a variety of transforms and derivative precipitation infomration, would be useful.

Like in the additive model, the forecasted GPP values for 2006 do not capture the drop in GPP values earlier than October.

Some peaks are forecasted to be much higher than they were. How is the 2016 spring peak forecast really high for GEDD2 whilst accurate for CHAL3?

Below we look at the peak in 2007 and then the peak in 2016.

### The peak of 2007
```{r gedd2007, echo = FALSE}
pg_by_yearmonth <- pggpp %>%
  group_by_key() %>%
  index_by(ym = yearmonth(times)) %>%
  summarise(monthlypg = sum(pg))
mean_monthly_pg <- pg_by_yearmonth %>% 
  as_tibble() %>%
  group_by(site, m = month(ym)) %>%
  summarise(mean_monthly_pg = mean(monthlypg))

est_resid_training %>%
  mutate(m = month(times)) %>%
  left_join(mean_monthly_pg, by = c("site", "m")) %>%
  filter(site %in% c("CHAL3", "GEDD2")) %>%
  filter_index("2005" ~ "2008") %>%
  pivot_longer(c(pg, .fitted), names_to = "valname") %>%
  as_tsibble(key = c(site,valname), index = times) %>%
  mutate(value = zoo::na.approx(value, rule = 1, na.rm = FALSE)) %>%
  #mutate(gpp = zoo::na.approx(gpp, rule = 1, na.rm = FALSE)) %>%
  ggplot() +
  facet_grid(rows = vars(valname), cols = vars(site), scales = "free_y") +
  geom_line(aes(x = times, y = value), na.rm = TRUE) +
  ggtitle("Forecasts and Observed GPP: training data") +
  geom_line(aes(x = times, y = mean_monthly_pg),
            function(x) dplyr::filter(x, valname == "pg"), #so only plots for pg facet
            lty = "dotted") + 
  geom_point(aes(x = times, y = gpp),
            function(x) dplyr::filter(x, valname == ".fitted"), #so only plots for pg facet
            col = "blue",
            na.rm = TRUE,
            size = 0.2) + 
  stat_summary_bin(aes(times, value),
         data = function(x) dplyr::filter(x, valname == "pg"), #so binning only applies to pg data
         colour = "blue",
         fun.y = sum,
         binwidth = 30,
         geom = "step")
```

In both sites the total rainfall before spring in 2007 looks about normal. 
The observed GPP fpr 2007 was similar to the GPP of a wet spring (2005)!
The forecast GPP was similar to the GPP of a dry years (2006).

Suggests:

1. There is a lowering of effect for any rain beyond middlish rain amounts. __remove high rain events from training?__
2. This could explain the forecasts too: forecast GPP not capturing this lower effect would mean the model fitting pulls all parameters closer to zero to avoid high estimates for large amounts of rain. This would be compounded to by trying to match GPP very close to 0.


### The peak of 2016
```{r gedd2016, echo = FALSE}
pg_by_yearmonth <- pggpp %>%
  group_by_key() %>%
  index_by(ym = yearmonth(times)) %>%
  summarise(monthlypg = sum(pg))
mean_monthly_pg <- pg_by_yearmonth %>% 
  as_tibble() %>%
  group_by(site, m = month(ym)) %>%
  summarise(mean_monthly_pg = mean(monthlypg))

est_resid_training %>%
  mutate(m = month(times)) %>%
  left_join(mean_monthly_pg, by = c("site", "m")) %>%
  filter(site %in% c("CHAL3", "GEDD2")) %>%
  filter_index("2010" ~ "2016") %>%
  pivot_longer(c(pg, .fitted), names_to = "valname") %>%
  as_tsibble(key = c(site,valname), index = times) %>%
  mutate(value = zoo::na.approx(value, rule = 1, na.rm = FALSE)) %>%
  #mutate(gpp = zoo::na.approx(gpp, rule = 1, na.rm = FALSE)) %>%
  ggplot() +
  facet_grid(rows = vars(valname), cols = vars(site), scales = "free_y") +
  geom_line(aes(x = times, y = value), na.rm = TRUE) +
  ggtitle("Forecasts and Observed GPP: training data") +
  geom_line(aes(x = times, y = mean_monthly_pg),
            function(x) dplyr::filter(x, valname == "pg"), #so only plots for pg facet
            lty = "dotted") + 
  geom_point(aes(x = times, y = gpp),
            function(x) dplyr::filter(x, valname == ".fitted"), #so only plots for pg facet
            col = "blue",
            na.rm = TRUE,
            size = 0.2) + 
  stat_summary_bin(aes(times, value),
         data = function(x) dplyr::filter(x, valname == "pg"), #so binning only applies to pg data
         colour = "blue",
         fun.y = sum,
         binwidth = 30,
         geom = "step")
```
Given that precipitation has relatively small coefficients in the model, how is the peak for GEDD2 spring 2016 so much higher than spring 2015 and previous years?

It is very strange that the observed GPP values do not increase in Spring 2016, even though a lot more precipitation fell around those months compared to previous years. This suggests there are physical thresholds in GPP (or in measuring/observing GPP) for each site.

a. how did the same model forecast a large peak for GEDD2 and a regular-sized peak for CHAL3 given that the precitation values are very similar? A: the prior year's GPP observation at CHAL3 are about 50% of the prior year's observed GPP at GEDD2.
b. Given the above, the amount of precipitation may not have impacted the forecast values at all.

2. there are sites that are all in the same pixel (so even though I've trained using 176 sites, really only 176/4 different locations at the resolution that this data is available).  __Suggest training only using the average of each farm__

The point b. can be tested by altering the observed gpp and reforcasting.


#### A different precipitation scenario for 2016
Replace the 2016 preciptiation with the 2015 precipitation

```{r alternativepg_2016}
alt_est_resid_training <- est_resid_training %>%
  mutate(lag1yr = lag(pg, n = 365.25)) %>%
  mutate(altpg = ifelse(year(times) == "2016", lag1yr, pg))

alt_est_resid_training %>%
  filter(site == "GEDD2") %>%
  filter_index("2015" ~ "2017") %>%
  pivot_longer(c(altpg, pg)) %>%
  ggplot() +
  geom_line(aes(x = times, y = value, lty = name, col = name)) +
  stat_summary_bin(aes(x = times, y = value, lty = name, col = name),
       fun.y = sum,
       binwidth = 30,
       geom = "step") +
  scale_color_viridis(discrete = TRUE) +
  ggtitle("Original and altered precipitation for 2016: GEDD2")
```

```{r forecastfromaltpg}
fit5_forecasts_l <- alt_est_resid_training %>% mutate(origpg = pg, pg = altpg) %>% select(-altpg) %>%
  filter(site %in% c("CHAL3", "GEDD2")) %>%
  group_by_key() %>%
  group_split() %>%
  lapply(function(x){
    new_tsbl = as_tsibble(x, key = NULL, index = "times")
    forecast(fit5, new_data = new_tsbl) %>%
      update_tsibble(key = site)
  })
fit5_forecasts <- fit5_forecasts_l %>%
  lapply(as_tibble) %>%
  (function(x) do.call(rbind, x)) %>%
  as_tsibble(key = "site", index = "times")

fit5_forecasts %>%
  filter(site %in% c("CHAL3", "GEDD2")) %>%
  filter_index("2016") %>%
  left_join(pggpp %>% select(times, site, gpp), by = c("times", "site"), suffix = c(".forecast", ".observed")) %>%
  mutate(interval = hilo(.distribution, 95)) %>%
  mutate(.fitted = zoo::na.approx(.fitted, rule = 1, na.rm = FALSE)) %>%
  ggplot() +
  facet_grid(rows = vars(site)) +
  geom_ribbon(aes(x = times, ymin = interval$.lower, ymax = interval$.upper), fill = "lightgrey") +
  geom_line(aes(x = times, y = gpp.forecast)) +
  geom_line(aes(x = times, y = .fitted), col = "orange") + 
  geom_point(aes(x = times, y = gpp.observed), col = "blue", size = 0.5) +
  ggtitle("Forecasts of GPP in 2016 under altered precipitation scenario.")
```

In the above yellow is the original forecast, black with grey ribbon is the new forecast that uses the *altered* precipitation, and blue dots are the observed GPP.
The precipitation has helped with the timing of the 2016 spring GPP for both CHAL3 and GEDD2, and it has caused the large spike in GEDD2 GPP forecast (yellow).

1. effect of high precipitation is too large. Consider: interaction terms, and removing very low GPP scenarios from the model.

### Is there something special about GEDD2 between 2010 and 2012?
There are multiple peaks in observed GPP for GEDD2 whilst the other sites have much smaller peaks, or no extra peaks at all.

```{r gedd2_multpeaks, echo = FALSE}
pggpp %>%
  filter(site %in% unique(pggpp$site)[c(1, 5, 9, 19, 54)]) %>%
  filter_index("2009" ~ "2013") %>%
  pivot_longer(c(pg, gpp), names_to = "valname") %>%
  as_tsibble(key = c(site, valname), index = times) %>%
  mutate(value = zoo::na.approx(value, rule = 1, na.rm = FALSE)) %>%
  ggplot() +
  facet_grid(rows = vars(valname), scale = "free_y") +
  geom_line(aes(x = times, y = value, col = site, lty = site), 
            lwd = 1.3,
            data = function(x) x %>% filter(valname == "gpp"),
            na.rm = TRUE) +
  stat_summary_bin(aes(x = times, y = value, col = site, lty = site),
     lwd = 1.3,
     data = function(x) x %>% filter(valname == "pg"),
     fun.y = sum,
     binwidth = 30,
     geom = "step") +
  scale_color_viridis_d(direction = -1) +
  ggtitle("Observed GPP and PG")
```

It looks like GEDD2 got a little bit more rain than the other 4 sites. But the GPP looks disproportionately larger!

I wonder if this is a symptom of the sites selected?

```{r gedd2_multpeaks2, echo = FALSE}
pggpp %>%
  filter(site %in% unique(pggpp$site)[seq(1, 173, by = 4)]) %>%
  filter_index("2009" ~ "2013") %>%
  pivot_longer(c(pg, gpp), names_to = "valname") %>%
  as_tsibble(key = c(site, valname), index = times) %>%
  mutate(value = zoo::na.approx(value, rule = 1, na.rm = FALSE)) %>%
  ggplot() +
  facet_grid(rows = vars(valname), scale = "free_y") +
  geom_line(aes(x = times, y = value, col = site), 
            lwd = 1.3,
            data = function(x) x %>% filter(valname == "gpp"),
            na.rm = TRUE) +
  stat_summary_bin(aes(x = times, y = value, col = site),
     lwd = 1.3,
     data = function(x) x %>% filter(valname == "pg"),
     fun.y = sum,
     binwidth = 30,
     geom = "step") +
  geom_line(aes(x = times, y = value),
            col = "red",
            lty = "dashed",
            lwd = 1,
            data = function(x) x %>% filter(valname == "gpp", site == "GEDD1"),
            na.rm = TRUE) +
  stat_summary_bin(aes(x = times, y = value, col = site),
     col = "red",
     lty = "dashed",
     lwd = 1,
     data = function(x) x %>% filter(valname == "pg", site == "GEDD1"),
     fun.y = sum,
     binwidth = 30,
     geom = "step") +
  scale_color_viridis_d(direction = -1) +
  ggtitle("Observed GPP and PG")
```

It looks like GEDD1 (red dashed) has unusally high GPP, but has rainfal that is only slightly higher than other sites.