---
title: "Look at recalculated LOO-PSIS of Past Models"
author: "Kassel Hingee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    collapsed: no
    number_sections: yes
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_root(rprojroot::has_file("DESCRIPTION")))
devtools::load_all(rprojroot::find_root(rprojroot::has_file("DESCRIPTION")))
knitr::opts_chunk$set(echo = FALSE)
library(tibble)
library(dplyr)
library(MCMCpack)
library(mclust)
library(corrplot)
library(coda)
library(runjags)
library(ggplot2)
library(patchwork)
```

## Data Import

```{r import_lpd}
indata_w1998 <- readRDS("./private/data/clean/7_2_4_input_data.rds")
indata_wo1998 <- readRDS("./private/data/clean/7_2_8_input_data.rds")

lpds_l <- c(readRDS("./tmpdata/7_2_4_lpds.rds"),
            readRDS("./tmpdata/7_2_5_lpds.rds"),
            readRDS("./tmpdata/7_2_6_lpds.rds"),
            sp2 = readRDS("./tmpdata/7_2_7_lpds.rds"),
            readRDS("./tmpdata/7_2_8_lpds.rds"))
names(lpds_l) <- gsub("sp2\\.","sp2_", names(lpds_l))
waics_l <- readRDS("./tmpdata/loo_many.rds")
setdiff(names(lpds_l), names(waics_l))
setdiff(names(waics_l), names(lpds_l))
```


## Log Posterior Density
```{r holdoutlpd}
melpd <- data.frame(Estimate = vapply(lpds_l, function(x) mean(x$lpds), FUN.VALUE = 3.3),
                    SE = vapply(lpds_l, function(x) return(sd(x$lpds) / sqrt(length(x$lpds))), FUN.VALUE = 3.3 ))

holdoutsizes <- vapply(lpds_l, function(x) length(x$lpds), FUN.VALUE = 3)
lpds_df_w1998 <- as_tibble(lapply(lpds_l[holdoutsizes == 245], function(x) x$lpds))
lpds_df_wo1998 <- as_tibble(lapply(lpds_l[holdoutsizes == 241], function(x) x$lpds))
lpds_df_wo1998[242:245, ] <- NA
lpds_df <- cbind(lpds_df_w1998, lpds_df_wo1998)


mean_2se <- function(x, mult = 2, na.rm = TRUE){
  if (na.rm == TRUE){x <- na.omit(x)}
  n <- length(x)
  xbar <- sum(x)/n
  sd <- sqrt(sum((x - xbar)^2)/(n - 1))
  return(c(ymin = xbar - mult * sd / sqrt(n),
           y = xbar,
           ymax = xbar + mult * sd / sqrt(n)))
}

lpds_df %>%
  as_tibble() %>%
  tidyr::pivot_longer(everything(), names_to = "model", values_to = "Site_lpd") %>%
  mutate(sp2 = grepl("sp2_", model)) %>%
  ggplot() +
  facet_grid(cols = vars(sp2), scales = "free") +
  # geom_violin(aes(x = Site_lpd, y = model)) +
  geom_violin(aes(x = Site_lpd, y = model),
              position = position_dodge(1),
              draw_quantiles = c(0.1, 0.25, 0.5, 0.75, 0.9)) +
  stat_summary(aes(x = Site_lpd, y = model),
               fun.data=mean_2se, fun.args = list(mult=2, na.rm = TRUE), geom="crossbar", width=0.2 ) +
  # stat_summary(aes(x = Site_lpd, y = model),
  #              fun = median, geom = "point", shape = 23, size = 2) +
  ggtitle("Log-Posterior Density of Holdout ModelSites")
```

Below is the lpd computed for each ModelSite for both the InSample and out of sample data.

```{r distributioninsampleoutsamplelpd}
insamplesizes <- vapply(waics_l, function(x) length(x$waic$pointwise[, "elpd_waic"]), FUN.VALUE = 3)

insamplelpds_df_w1998 <- do.call(cbind, lapply(waics_l[insamplesizes == 2101], function(x) x$waic$pointwise[, "elpd_waic"])) %>%
  as_tibble() %>%
  mutate(SurveySiteId = indata_w1998$insampledata$Xocc[, "SurveySiteId", drop = TRUE],
         SurveyYear = indata_w1998$insampledata$Xocc[, "SurveyYear", drop = TRUE]) %>%
  tidyr::pivot_longer(c(-SurveySiteId, -SurveyYear), names_to = "model", values_to = "site_lpd")
insamplelpds_df_wo1998 <- do.call(cbind, lapply(waics_l[insamplesizes == 2059], function(x) x$waic$pointwise[, "elpd_waic"])) %>%
  as_tibble() %>%
  mutate(SurveySiteId = indata_wo1998$insampledata$Xocc[, "SurveySiteId", drop = TRUE],
         SurveyYear = indata_wo1998$insampledata$Xocc[, "SurveyYear", drop = TRUE])
insamplelpds_df_wo1998[2060:2101, ] <- NA 
insamplelpds_df_wo1998 <- insamplelpds_df_wo1998 %>%
  tidyr::pivot_longer(c(-SurveySiteId, -SurveyYear), names_to = "model", values_to = "site_lpd")
insamplelpds_df <- bind_rows(insamplelpds_df_w1998, insamplelpds_df_wo1998) %>%
  mutate(InSample = TRUE)

df <- lpds_df %>% as_tibble() %>% 
  tidyr::pivot_longer(everything(), names_to = "model", values_to = "site_lpd") %>%
  mutate(SurveySiteId = NA, SurveyYear = NA) %>% 
  mutate(InSample = FALSE)

df <- bind_rows(insamplelpds_df, df)


df %>%
  as_tibble() %>%
  mutate(TwoSpeciesOnly = grepl("sp2_", model)) %>%
  filter(TwoSpeciesOnly != TRUE) %>%
  ggplot() +
  facet_grid(rows = vars(model), scales = "free", switch = "y") +
  geom_violin(aes(x = site_lpd, y = InSample, col = InSample), position = position_dodge(1)) +
  geom_point(aes(x = site_lpd, y = InSample, col = InSample), position = position_dodge(1)) +
  stat_summary(aes(x = site_lpd, y = InSample, group = InSample),
               fun.data=mean_2se, fun.args = list(mult=2), geom="crossbar", width=0.2 ) +
  theme(strip.text.y.left = element_text(angle = 0)) +
  ggtitle("Log Posterior Density of ModelSites")

df %>%
  as_tibble() %>%
  mutate(TwoSpeciesOnly = grepl("sp2_", model)) %>%
  filter(TwoSpeciesOnly == TRUE) %>%
  ggplot() +
  facet_grid(rows = vars(model), scales = "free", switch = "y") +
  geom_violin(aes(x = site_lpd, y = InSample, group = InSample),
              position = position_dodge(1),
              draw_quantiles = c(0.1, 0.25, 0.5, 0.75, 0.9)) +
  geom_point(aes(x = site_lpd, y = InSample, col = InSample), position = position_dodge(1)) +
  stat_summary(aes(x = site_lpd, y = InSample, group = InSample),
               fun.data=mean_2se, fun.args = list(mult=2), geom="crossbar", width=0.2 ) +
  theme(strip.text.y.left = element_text(angle = 0)) +
  ggtitle("Log Posterior Density of ModelSites")
```

How can the uncorrected lpd of insample data for the gppmean model be *worse* than the intercepts only model?
Similarly for msnm_time, and many other models. MCMC must not be maximising lpd.
But if things were set up well I would expect it do a good job.

Is it due to outliers?

```{r justtwomodels}
lowvals <- df %>%
  filter(InSample) %>%
  filter(model %in% "intercepts_only") %>%
  filter(site_lpd < -100) %>%
  mutate(low_lpd_intercepts_only = TRUE)
lowvals[, c("SurveySiteId", "SurveyYear", "site_lpd")] %>% arrange(site_lpd)

left_join(df, lowvals[, c("SurveySiteId", "SurveyYear", "low_lpd_intercepts_only")],
          by = c("SurveySiteId", "SurveyYear"))  %>%
  mutate(TwoSpeciesOnly = grepl("sp2_", model)) %>%
  filter(TwoSpeciesOnly != TRUE) %>%
  filter(model %in% c("intercepts_only", "msnm")) %>%
  ggplot() +
  facet_grid(rows = vars(model), scales = "free", switch = "y") +
  geom_violin(aes(x = site_lpd, y = InSample, group = InSample),
              position = position_dodge(1),
              draw_quantiles = c(0.1, 0.25, 0.5, 0.75, 0.9)) +
  geom_point(aes(x = site_lpd, y = InSample, group = InSample),
             data = function(x) x %>% filter(!is.finite(low_lpd_intercepts_only)),
             position = position_jitter(height = 0.3),
             alpha = 0.2,
             shape = 4) +
  geom_point(aes(x = site_lpd, y = InSample, group = InSample, shape = as.factor(SurveySiteId), col = SurveyYear),
             data = function(x) x %>% filter(low_lpd_intercepts_only),
             position = position_jitter(height = 0.1)) +
  scale_color_viridis_c() +
  scale_shape_manual(values = 1:20) +
  stat_summary(aes(x = site_lpd, y = InSample, group = InSample),
               fun.data=mean_2se, fun.args = list(mult=2), geom="crossbar", width=0.2 ) +
  theme(strip.text.y.left = element_text(angle = 0)) +
  ggtitle("Log Posterior Density of ModelSites")
```

The three worst Model Sites are the same for both the intercepts and msnm model.


```{r worst3sites_plottedforallmodels}
left_join(df, lowvals[, c("SurveySiteId", "SurveyYear", "low_lpd_intercepts_only")],
          by = c("SurveySiteId", "SurveyYear"))  %>%
  mutate(TwoSpeciesOnly = grepl("sp2_", model)) %>%
  filter(TwoSpeciesOnly != TRUE) %>%
  filter(InSample) %>%
  ggplot() +
  facet_grid(rows = vars(model), scales = "free", switch = "y") +
  geom_violin(aes(x = site_lpd, y = InSample, group = InSample),
              position = position_dodge(1),
              draw_quantiles = c(0.1, 0.25, 0.5, 0.75, 0.9)) +
  geom_point(aes(x = site_lpd, y = InSample, group = InSample, shape = as.factor(SurveySiteId), col = SurveyYear),
             data = function(x) x %>% filter(low_lpd_intercepts_only),
             position = position_jitter(height = 0.1)) +
  scale_color_viridis_c() +
  scale_shape_manual(values = 1:20) +
  stat_summary(aes(x = site_lpd, y = InSample, group = InSample),
               fun.data=mean_2se, fun.args = list(mult=2), geom="crossbar", width=0.2 ) +
  theme(strip.text.y.left = element_text(angle = 0)) +
  ggtitle("Log Posterior Density of In-Sample ModelSites")
```

These three sites are poor for *all* the models. 
Do these sites have high p_waic or k value?
What is so unusual about these sites? It is hard to tell from the visit data:

```{r worst3sitsobs}
inner_join(indata_w1998$insampledata$yXobs, lowvals[, c("SurveySiteId", "SurveyYear", "low_lpd_intercepts_only")],
          by = c("SurveySiteId", "SurveyYear")) %>%
  dplyr::select(SurveySiteId, SurveyYear, all_of(indata_w1998$species))
```

These outliers don't indicate why WAIC or lpd is not discriminating between models though.


```{r waics}
waics_average_elpd_per_point <- lapply(waics_l, function(x) x$waic$estimates["elpd_waic", ]/nrow(x$waic$pointwise))
waics_average_elpd_per_point <- simplify2array(waics_average_elpd_per_point)
waics_average_elpd_per_point <- t(waics_average_elpd_per_point) %>% as_tibble(rownames = "Model")
waics_average_elpd_per_point$type = "WAIC"

loo_average_elpd_per_point <- lapply(waics_l, function(x) x$loo$estimates["elpd_loo", ]/nrow(x$waic$pointwise))
loo_average_elpd_per_point <- simplify2array(loo_average_elpd_per_point)
loo_average_elpd_per_point <- t(loo_average_elpd_per_point) %>% as_tibble(rownames = "Model")
loo_average_elpd_per_point$type = "LOO-PSIS"

melpd$type <- "Holdout"
melpd <- as_tibble(melpd, rownames = "Model")

df <- bind_rows(waics_average_elpd_per_point, loo_average_elpd_per_point, melpd)

df %>%
  mutate(TwoSpeciesOnly = grepl("sp2_", Model)) %>%
  ggplot() +
  facet_grid(cols = vars(TwoSpeciesOnly), scales = "free") +
  geom_errorbar(aes(x = Model, ymax = Estimate +  2*SE, ymin = Estimate - 2*SE, col = type, lty = type)) +
  geom_point(aes(x = Model, y = Estimate, col = type), position = position_jitter(width = 0.1)) +
  coord_flip() +
  ggtitle("Estimates of Log Posterior Density for a new ModelSite")
```

There is only one model where the new (better) LOO-PSIS estimates are different to the WAIC estimates.
